---
title: "simpleHydrograph and hydroStats"
author: "Kevin Shook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(MESHr)
library(ggplot2)
MESH_streamflows <- MESH_streamflows
```

The simpleHydrograph function is intended to plot hydrographs from MESH output alone,
i.e. without additional data being supplied from another source, such as WSC data.
The function produces a **ggplot2** object, which can then be manipulated to customise
the plot. 

The function requires a data frame of MESH flows. The first column must be an
**R** date (named DATE) or a **POSIXct** datetime (named DATETIME). The easiest
way to produce this data frame is by reading in the data using the function
**readOutputTimeseriesCSV**. Functions will be added to read the output directly
from the netCDF output.

Note that you can subset the data to restrict the range plotted.

The simplest plot shows all of the data in a single un-facetted graph. The stations
are denoted by colour, and the type of flow (i.e. measured or simulated), by 
line type. By default, the stations are labelled by their MESH numbers.
```{r, fig.width = 5, fig.height = 3}
simpleHydrograph(MESH_streamflows)
```

You can change the station names by specifying them in a vector.
```{r, fig.width = 5, fig.height = 3}
station_names <- c("Station 1", "Station 2")
simpleHydrograph(MESH_streamflows, stationNames = station_names)
```


The function encourages the use of facetting to produce more readable plots. When facetting by station,
it's a good idea to turn off the coloring of plots by station, by setting `byStation = FALSE`. You can
change the x-axis scale and/or labels. Note that this example forces the facets to be
on separate rows, so that the x-axes are as long as possible. Because this requires other functions
in ggplot2, you should load it before proceeding.

```{r, fig.width = 5, fig.height = 3}
p <- simpleHydrograph(MESH_streamflows, stationNames = station_names, byStation = FALSE)
library(ggplot2)
p <- p + facet_wrap(~station, nrow = 2) + scale_x_date(date_labels = "%Y")
p
```

Because the magnitudes of the flows are so different for the stations, you can use 
separate y scales:
```{r, fig.width = 5, fig.height = 3}
p <- p + facet_wrap(~station, nrow = 2, scales = "free_y") 
p
```

Alternatively, you can facet the plots by the year. Note that this option adds a
variable to the plot called `YEAR`. The uppercase variable name is in keeping with the 
other MESH variables.

```{r, fig.width = 5, fig.height = 3}
p <- simpleHydrograph(MESH_streamflows, stationNames = station_names, byYear = TRUE)
p <- p + facet_wrap(~YEAR, scales = "free_y")
p
```

You can also create gridded facets by year and station.
```{r, fig.width = 5, fig.height = 6}
p <- simpleHydrograph(MESH_streamflows, stationNames = station_names, byStation = FALSE, byYear = TRUE)
p <- p + facet_grid(YEAR~station, scales = "free_y")
p
```

The default colors are a bit bland - you can easily change them. In this example,
the y-axis labels are rotated, so they show up better.
```{r, fig.width = 5, fig.height = 6}
plotcols <- c("red", "blue")
p <- p + scale_colour_manual(values = plotcols) +
  theme(strip.text.y = element_text(angle = 0))
p
```

You can plot only the measured or the simulated values. This this case, it's a good idea to suppress the flow type legend.
```{r, fig.width = 5, fig.height = 6}
p <- simpleHydrograph(MESH_streamflows, stationNames = station_names, 
                      byStation = FALSE, byYear = TRUE, meas = FALSE) +
  facet_grid(YEAR~station, scales = "free_y") +
  scale_colour_manual(values = plotcols) +
  theme(strip.text.y = element_text(angle = 0)) +
  theme(legend.position = "none")

p
```

You can use the function **hydroStats** to calculate the goodness of fit of your 
modelled hydrograph. 
```{r}
stats <- hydroStats(MESH_streamflows, stationNames = station_names)
stats
```

Once the stats have been calculated, they can be added to a hydrograph. You can add the station names
using the **ggplot2** function **annotate**
To select the stat you want, specify its variable name. You can combine values together,
or add station names.
```{r}
NSE <- paste(stats$station, "NSE =", stats$NSE, sep = " ")
NSE
```

To add both values as a single string (which is required to annotate a single graph), you can paste them together
```{r}
all_NSE <- paste(NSE, collapse = ", ")
all_NSE
```

Then use annotate to add the string to the graph. Note that you have to specify the
location of the string, relative to the axes. Because the x-axis uses dates,
you have to specify the location as a date.
```{r, fig.width = 5, fig.height = 3}
p1 <- simpleHydrograph(MESH_streamflows, stationNames = station_names, byStation = TRUE) +
  scale_x_date(date_labels = "%Y")

dateLoc <- as.Date("2005-01-01", format = "%Y-%m-%d")

p2 <- p1 + annotate("text", x = dateLoc, y = 1500, label = all_NSE, hjust = "left")
p2
```

Since the text doesn't fit very well, you can wrap it before adding it to the graph.
```{r, fig.width = 5, fig.height = 3}
wrapped_NSE <- paste(NSE, collapse = "\n")
p3 <- p1 + annotate("text", x = dateLoc, y = 1500, label = wrapped_NSE, hjust = "left")
p3
```

You can also apply the vector of NSE values to plot facets.
```{r, fig.width = 5, fig.height = 3}
p4 <- simpleHydrograph(MESH_streamflows, stationNames = station_names, byStation = FALSE)
p4 <- p4 + facet_wrap(~station, nrow = 2) + scale_x_date(date_labels = "%Y")
p4 <- p4 + annotate("text", x = dateLoc, y = 1500, label = NSE, hjust = "left")
p4
```

You can specify that the modelled data consist of both caibration 
and verification data, by specifying the beginning and/or the end
dates of the calibration series. This adds the variable ***period**
to the plot, which can then be used for facetting. The value of
the variable will be **Calibration** during the calibration period
and **Validation** during the rest.

```{r, fig.width = 5, fig.height = 3}
calEnd <- "2010-01-01"
p5 <- simpleHydrograph(MESH_streamflows, stationNames = station_names, 
                       byStation = FALSE, calEnd = calEnd)
p5 <- p5 + facet_grid(station~period) + scale_x_date(date_labels = "%Y")
p5
```

You can easily change the scales (x, y, and colour) to make them more attractive.
```{r, fig.width = 5, fig.height = 3}
plotcols <- c("red", "blue")
p6 <- p5 + facet_grid(station~period, scales = "free") + 
  scale_x_date(date_labels = "%Y") +
  scale_colour_manual(values = plotcols)
p6
```

The **hydroStats** function also works with separate calibration and validation 
periods.
```{r}
stats2 <- hydroStats(MESH_streamflows, stationNames = station_names, calEnd = calEnd)
stats2
```

You can add the stats by period to the plot.
```{r, fig.width = 5, fig.height = 3}
facetted_NSE <- paste("NSE = ", stats2$NSE[order(stats2$station)])
p7 <- p5 + facet_grid(station~period) + 
  scale_x_date(date_labels = "%Y") +
  scale_colour_manual(values = plotcols) + 
  annotate("text", x = dateLoc, y = 1500, label = facetted_NSE, hjust = "left")
p7
```

