---
title: "simpleHydrograph"
author: "Kevin Shook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(MESHr)
library(ggplot2)
MESH_streamflows <- MESH_streamflows
```

The simpleHydrograph function is intended to plot hydrographs from MESH output alone,
i.e. without additional data being supplied from another source, such as WSC data.
The function produces a **ggplot2** object, which can then be manipulated to customise
the plot. 

The function requires a data frame of MESH flows. The first column must be a
**R** date (named DATE) or a **POSIXct** datetime (named DATETIME). The easiest
way to produce this data frame is by reading in the data using the function
**readOutputTimeseriesCSV**. Functions will be added to read the output directly
from the netCDF output.

Note that you can subset the data to restrict the range plotted.

The simplest plot shows all of the data in a single un-facetted graph. The stations
are denoted by colour, and the type of flow (i.e. measured or simulated), by 
line type. By default, the stations are labelled by their MESH numbers.
```{r, fig.width = 5, fig.height = 3}
simpleHydrograph(MESH_streamflows)
```

You can change the station names by specifying them in a vector.
```{r, fig.width = 5, fig.height = 3}
station_names <- c("Station 1", "Station 2")
simpleHydrograph(MESH_streamflows, stationNames = station_names)
```


The function encourages the use of facetting to produce more readable plots. When facetting by station,
it's a good idea to turn off the coloring of plots by station, by setting `byStation = FALSE`. You can
change the x-axis scale and/or labels. Note that this example forces the facets to be
on separate rows, so that the x-axes are as long as possible.

```{r}
p <- simpleHydrograph(MESH_streamflows, stationNames = station_names, byStation = FALSE)
p <- p + facet_wrap(~station, nrow = 2) + scale_x_date(date_labels = "%Y")
p
```

Because the magnitudes of the flows are so different for the stations, you can use 
separate y scales:
```{r, fig.width = 5, fig.height = 3}
p <- p + facet_wrap(~station, nrow = 2, scales = "free_y") 
p
```

Alternatively, you can facet the plots by the year. Note that this option adds a
variable to the plot called `YEAR`. The uppercase variable name is in keeping with the 
other MESH variables.

```{r, fig.width = 5, fig.height = 3}
p <- simpleHydrograph(MESH_streamflows, stationNames = station_names, byYear = TRUE)
p <- p + facet_wrap(~YEAR, scales = "free_y")
p
```

You can also create gridded facets by year and station.
```{r, fig.width = 5, fig.height = 6}
p <- simpleHydrograph(MESH_streamflows, stationNames = station_names, byStation = FALSE, byYear = TRUE)
p <- p + facet_grid(YEAR~station, scales = "free_y")
p
```

The default colors are a bit bland - you can easily change them. In this example,
the y-axis labels are rotated, so they show up better.
```{r, fig.width = 5, fig.height = 6}
plotcols <- c("red", "blue")
p <- p + scale_colour_manual(values = plotcols) +
  theme(strip.text.y = element_text(angle = 0))
p
```

You can plot only the measured or the simulated values. This this case, it's a good idea to suppress the flow type legend.
```{r, fig.width = 5, fig.height = 6}
p <- simpleHydrograph(MESH_streamflows, stationNames = station_names, 
                      byStation = FALSE, byYear = TRUE, meas = FALSE) +
  facet_grid(YEAR~station, scales = "free_y") +
  scale_colour_manual(values = plotcols) +
  theme(strip.text.y = element_text(angle = 0)) +
  theme(legend.position = "none")

p
```

